# Biomovement Data and Exercise Quality Prediction
## pmccullo February 2015
### Coursera Reproducible Machine Learning
========================================================

**Synopsis**

This project seeks to build a prediciton model for how well exercises are 
performed based on data from http://groupware.les.inf.puc-rio.br/har

This project uses the Caret package in R to perform machine learning.

**Data Processing**

Load packages to be used:

```{r message = FALSE}
 library(caret); library(RWeka)
 set.seed(1000)
```

*Download and Load Data*

```{r}
  filename <- c("./pml-training.csv")
  url <- c("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
  download.file(url,destfile = filename, method = "curl")
  originaldata <- read.csv(filename)

  filename <- c("./pml-testing.csv")
  url <- c("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
  download.file(url,destfile = filename, method = "curl")
  testingdata <- read.csv(filename)
  
```

**Exploratory Analysis**

The data provided is a very large dataset:

```{r}
ncol(originaldata)
nrow(originaldata)
ncol(originaldata)*nrow(originaldata)
```

There are 5 classes of performance:
- Class A = well executed
- Class B = throwing the elbows to the front
- Class C = lifting dumbbell halfway
- Class D = lowering dummbbell halfwway
- Class E = throwing hips to the front

There are 4 sensors and 37 measurements per sensor:
- Belt sensor is located at the belt buckle (col 8-45)
- Arm sensor is located on upper arm (right above the elbow) - (col 46-83)
- Dumbbell sensor is located on the dumbbell (col 84-121)
- Forearm sensor is located near the wrist (col 122-159)

There are a number of columns that the researchers have added calculations into
that either show up at blanks or as NAs for the majority of the entries. This 
analysis will ignore those columns.

```{r}
calccol <- c(12:36,50:59,69:83,87:101,103:112,125:139,141:150)
trimdata <- originaldata[,-calccol]
```

To trim that list down, we will use the nearZeroVar function to identify variables
that we can ignore. This removed 1 column.

```{r}
  nsv <- nearZeroVar(trimdata, saveMetrics = FALSE)
  trimdata2 <- trimdata[,-nsv]
```

Taking out the label columns to only include predictors and outcome. This step
removes the lable columns (timestamps and similar information).

```{r}
  usedata <- trimdata2[,-c(1:6)]
```

I decided to attempt a tree based analysis as I believe that it would provide a
potentially quick way of providing feedback when errors occur in the movement.

Using a 6 fold cross validation methodology, as a way to both protect against
overfitting a single set of data as well as to provide insight into the potential
out of sample error rate. The 6 fold approach was chosen to balance the time and
the value add of additional sampling.

```{r}
fitControl <- trainControl(method = "cv", number = 6)
modfit <- train(classe ~ ., data = usedata, method = "LMT",
                trControl = fitControl)
modfit$resample
```

The accuracy rates all exceed 98%. This indicates that for this sample, this model
is fairly accurate. My expectation is that the out of sample rate for similarily 
executed data will be exceed 90%. I believe that this model will 99% of the time
guess greater than 14 of the 20 test cases (assuming they use the similar 
experiment structure).

However, this exercise was extremely regulated compared to normal weight lifting. 
The weight involved was on the order or 10x smaller than one would anticipate 
using. If this system was implemented outside of the lab, I would expect the accuracy
to drop to closer to 50% with a margin of +/- 25%. 
